\documentclass[12pt]{pset}
%\usepackage{add-copyright}

\title{Problem Set 6}
\course{Piecewise-Linear Topology}
\author{Jim Fowler}
\date{Summer 2010}

\newcommand{\defnword}[1]{\textbf{#1}}

\usepackage{stmaryrd}
\newcommand{\boundary}{\partial}
\newcommand{\collapses}{\searrow}
\newcommand{\expands}{\nearrow}
\newcommand{\she}{\ssearrow\nnearrow}
\newcommand{\join}{\ast}
\newcommand{\subdivided}{\triangleleft}
\geometry{margin=1in}%,top=0.5in,bottom=0.5in}
\DeclareMathOperator{\st}{st}
\DeclareMathOperator{\vertices}{vert}
\DeclareMathOperator{\lk}{lk}
\DeclareMathOperator{\cl}{cl}
\DeclareMathOperator{\interior}{int}
\newcommand{\fullsubcomplex}{\Subset}

\begin{document}
\maketitle

\subsubsection*{So many exercises.} There are many things that we can
prove now that we have our new definitions of polyhedra and cell
complexes.  These exercises are mostly from Rourke and Sanderson's
textbook, \textit{Introduction to Piecewise-Linear Topology}.

\parskip 0.5\baselineskip
\parindent 0pt

\subsection*{Polyhedra}

\begin{problem}
  Is the intersection of finitely many polyhedra a polyhedron?
\end{problem}

\begin{problem}
  Is the union of finitely many polyhedra a polyhedron?
\end{problem}

\begin{problem}
  Suppose $P \subset \R^n$ is a compact polyhedron; let $v$ be a point in
  $\R^{n+1} - \R^n$, and show that $v \join P$ is a polyhedron.
\end{problem}

\begin{requiredproblem}
  Show that the composition of PL maps is again a PL map.
\end{requiredproblem}

\begin{requiredproblem}
  Is the product of an $n$-manifold (with boundary) and an
  $m$-manifold (with boundary) an $nm$-manifold?
\end{requiredproblem}

\begin{problem}
  Show that a compact polyhedron is the finite union of simplexes.
  Use this to show that the PL image of a compact polyhedron is a
  polyhedron.
\end{problem}

\begin{problem}
  Define the ``dimension'' of a polyhedron, and show that it is
  invariant under PL homeomorphism.
\end{problem}

\begin{problem}
  If $A \join B \cong S^n$, is it the case that $A$ and $B$ are spheres?
\end{problem}

\subsection*{Cell complexes}

\begin{requiredproblem}
 Is the intersection of two cells also a cell?
\end{requiredproblem}

\begin{problem}
 Is the intersection of two cell complexes also a cell complex?
\end{problem}

\begin{problem}
  Is the cone of a cell complex itself a cell complex?
\end{problem}

\begin{requiredproblem}
  Show that the product of cell complexes is a cell complex.
\end{requiredproblem}

\subsection*{Surfaces}

\begin{requiredproblem}
 The cube $I^4$ can be regarded as a cell complex; find a subcomplex
 of $I^4$ consisting of squares which is PL homeomorphic to
 the torus $T^2$.
\end{requiredproblem}

\begin{problem}
  Find a subcomplex of $I^5$ (which will consist of squares) which is
  PL homeomorphic to a genus five surface.
\end{problem}

\subsection*{Simplicial complexes}

These problems are \textbf{not} about abstract simplicial complexes,
but rather, the sort of simplicial complexes that sit in $\R^n$.

\begin{problem}
  Show that a cell complex can be subdivided to a simplicial complex.
\end{problem}

\begin{problem}
  Suppose $f : K \to L$ is a simplicial map, and $L' \subdivided L$.
  Find a subdivision $K' \subdivided K$ so that $f : K' \to L'$ is
  simplicial.
\end{problem}

\begin{problem}
  If $f : |K| \to |K|$ is a PL map, and $f^2$ is the identity, is
  there a subdivision $K' \subdivided K$ so that $f : K' \to K'$ is
  simplicial?
\end{problem}

\begin{problem}
  Let $K$ and $L$ be simplicial complexes, and $f : |K| \to |L|$ a PL
  homeomorphism.  For $a \in K$, show that there is a PL homeomorphism
  $|\lk(a,K)| \cong |\lk(f(a),L)|$.
\end{problem}

\begin{problem}
  Show that if $K$ is a simplicial complex, then $|K|$ is an
  $n$-manifold without boundary if and only if $|\lk(v,K)| \cong
  S^{n-1}$ for each $v \in K$.
\end{problem}

\begin{problem}
  Show that $|K| \times \R$ is a PL manifold if and only if $|K|$ is a manifold.
\end{problem}

\end{document}

