---
title: Aspherical manifolds that cannot be triangulated
status: published
arxiv: 1304.3730v2
year: 2013
mr: MR3159970
summary: "Kirby and Siebenmann showed that there are manifolds that do not admit PL structures, and yet the possibility remained that all manifolds could be triangulated, meaning that for every manifold $M$, there is a simplicial complex $K$ so that the geometric realization of $K$ is homeomorphic to $M$, but of course the simplicial complex $K$ is not a PL triangulation, meaning the links are not spheres.  Freedman showed that there are 4-manifolds that cannot be triangulated.  Davis and Januszkiewicz applied a hyperbolization procedure to Freedman's 4-manifolds to get closed aspherical 4-manifolds that cannot be triangulated.  What about higher dimensions?"
joint:
 - Michael W. Davis
 - Jean-François Lafont
---

Kirby and Siebenmann showed that there are manifolds that do not admit PL structures \cite{MR242166}, and yet the possibility remained that all manifolds could be triangulated, meaning that for every manifold $M$, there is a simplicial complex $K$ so that the geometric realization of $K$ is homeomorphic to $M$, but of course the simplicial complex $K$ is not a PL triangulation, meaning the links are not spheres.

Freedman showed that there are 4-manifolds that cannot be triangulated \cite{MR679066}.  Davis and Januszkiewicz applied a hyperbolization procedure to Freedman's 4-manifolds to get closed aspherical 4-manifolds that cannot be triangulated \cite{MR1131435}.  What about higher dimensions?

In the late 1970s Galewski and Stern \cite{MR558395} and independently, Matumoto, showed that non-triangulable manifolds exist in all dimensions $> 4$ if and only if homology $3$-spheres with certain properties do not exist. In \cite{MR3122335}, Manolescu showed that there were no such homology $3$-spheres, and hence non-triangulable manifolds exist in every dimension $>4$.

Applying a hyperbolization technique to the Galewski-Stern manifolds shows the following.
\begin{theorem}[Davis--Lafont--Fowler]
  Let $n > 5$.  There exists a closed aspherical $n$-manifold which
  cannot be triangulated.
\end{theorem}
However, the following question remains open: Do there exist closed aspherical $5$-manifolds that cannot be triangulated?
