\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsmath}

\newtheorem*{example}{Example}
\newtheorem*{thm}{Theorem}

\title{Lecture 3: Vector products}
\author{Math 195 Section 91}
\date{Friday June 26, 2009}

\begin{document}
\maketitle

Goal: section 13.3 and 13.4

\subsection*{Dot product}

definition

properties:
\begin{itemize}
\item $v \cdot v = |v|^2$
\item $v \cdot (w + u) = v \cdot w + v \cdot u$
\item $0 \cdot v = 0$.
\item $v \cdot w = w \cdot v$
\item $(\alpha v) \cdot w = \alpha \cdot (v \cdot w) = v \cdot (\alpha w)$.
\end{itemize}

use dot product to find length.  define unit vector.

what does it measure: the angle $v \cdot w = |v| \cdot |w| \cdot \cos \theta$.

how to find the angle

prop: vectors $v$ and $w$ are orthognoal if $v \cdot w = 0$.

especially fun to find angle with $i, j, k$.

\subsection{Projection}

project $w$ onto $v$.

use projection to write $w$ as a combination of $u$ and $v$.

\subsection{higher dimensions!}

we can do dot products in any dimension!  can you compute the angle
between two four dimensional vectors?

\subsection{Cross product}

cross product is something special about 3-dimensions---takes two vectors and gives another vector

define cross product using the ``circle $ijk$''.  mention that the book uses determinants, but that we won't do this.

theorem: $a \times b$ is perpendicular to $a$ and $b$.  proof?  dot products!

theorem: $|a \times b| = |a| \cdot |b| \cdot \sin \theta$.  This is the same asthe area of the parallelogram.

interactively deduce the varous properties of cross products

commutative?  associative?  

\end{document}
