\documentclass[12pt]{pset}
%\usepackage{add-copyright}

\title{Problem Set 8}
\course{Piecewise-Linear Topology}
\author{Jim Fowler}
\date{Summer 2010}

\newcommand{\defnword}[1]{\textbf{#1}}

\usepackage{stmaryrd}
\newcommand{\boundary}{\partial}
\newcommand{\collapses}{\searrow}
\newcommand{\expands}{\nearrow}
\newcommand{\she}{\ssearrow\nnearrow}
\newcommand{\join}{\ast}
\newcommand{\subdivided}{\triangleleft}
\geometry{margin=1in}%,top=0.5in,bottom=0.5in}
\DeclareMathOperator{\st}{st}
\DeclareMathOperator{\vertices}{vert}
\DeclareMathOperator{\lk}{lk}
\DeclareMathOperator{\cl}{cl}
\DeclareMathOperator{\interior}{int}
\newcommand{\fullsubcomplex}{\Subset}

\begin{document}
\maketitle

\subsubsection*{More things collapse.} Here are a few more problems
on simplicial collapse.

%\parskip 0.25\baselineskip
%\parindent 0pt

\begin{requiredproblem}
  Show that any 1-dimensional complex is simple homotopy equivalent to
  $S^1 \vee \cdots \vee S^1$, i.e., the union of circles along one point.
\end{requiredproblem}

\begin{requiredproblem}
  Find a 2-complex which is not simple homotopy equivalent to a
  1-dimensional complex (and, of course, provide a proof of the fact
  that your 2-complex satisfies this property).
\end{requiredproblem}

\begin{problem}
  Find a proper embedding $B^1 \subset B^3$ so that $B^3$ does not
  collapse to $B^1$.  Here, proper means that the boundary is sent to
  the boundary, and the interior to the interior.
\end{problem}

\subsection*{Slice knots}

\begin{problem}
 The trefoil knot $S^1$ sits in $S^3$; find a disk $D^2$
 properly embedded in $B^4$, so that the boundary of $D^2$ is
 the trefoil knot in $S^3$.  Here, \textbf{properly} means that the
 boundary is sent to the boundary, and the interior to the interior.
\end{problem}

\begin{problem}
  Find a surface $\Sigma^2$ properly and \textit{locally flatly}
  embedded in $B^4$, so that the boundary of $\Sigma^2$ is the trefoil
  knot in $S^3$.  Here, \textbf{locally flatly} means that the link of
  a vertex of $\Sigma^2$ in the pair $(B^4,\Sigma^2)$ is an unknotted
  $(S^3,S^1)$.
\end{problem}

\begin{problem}
  Do the same for the figure eight knot.  In these cases, what is the
  genus of the surface that you used?  Can you reduce the genus?  How
  low can the genus go?
\end{problem}


\end{document}

