\documentclass[12pt]{handout}

\course{Math 765}
\title{Take-Home Final Exam}
\author{Jim Fowler}
\date{Winter 2011}

\newcommand{\CC}{\mathbb{C}}
\newcommand{\CP}{\mathbb{C}P}
\newcommand{\RP}{\mathbb{R}P}

\usepackage{enumerate}
\usepackage{nopageno}
\usepackage{hyperref}
\usepackage{tikz}

\DeclareMathOperator{\Vect}{Vect}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SO}{SO}
\DeclareMathOperator{\so}{\mathfrak{so}}
\DeclareMathOperator{\codim}{codim}
\DeclareMathOperator{\Mat}{Mat}
\DeclareMathOperator{\SymMat}{SymMat}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\Sym}{Sym}
\newcommand{\Proj}{\mathbb{P}}
\newcommand{\LL}{\mathcal{L}}
\DeclareMathOperator{\Div}{div}
\DeclareMathOperator{\grad}{grad}

\newcommand{\HH}{H}
\newcommand{\hh}{\mathfrak{h}}

\DeclareMathOperator{\Imm}{Imm}
\DeclareMathOperator{\Mono}{Mono}
\usepackage{amsthm}
\newtheorem*{theorem*}{Theorem}

\begin{document}
\maketitle

\noindent\textbf{The final exam.}  The course is ended, but there is
still one task that remains---this, the take-home final exam.  Please
feel free to email me with questions at \texttt{fowler@math.osu.edu},
and feel free to discuss the problems with your friends---but write up
your own solutions.
\textit{The final exam below should be handed in 100 hours from now---on Tuesday, March 15,
  2011, at 5:30\textsc{p.m.}}

\subsection*{Problem 1}

Consider $G_2(\R^4)$, the Grassmannian of planes in $\R^4$ (that is, a
point in $G_2(\R^4)$ corresponds to a subspace $\R^2 \hookrightarrow
\R^4$).  Show that the map
$$
\bigsqcup_{\stackrel{V \subset \R^4}{\dim V = 2}} V \to G_2(\R^4)
$$
sending a vector in $V$ to the point $[V] \in G_2(\R^4)$ is a smooth
vector bundle.

\subsection*{Problem 2}

Let $M$ be a closed smooth $n$-manifold without boundary; let $g$ be a
Riemannian metric on $M$; the Laplacian $\Delta : C^\infty(M) \to
C^\infty(M)$ is given by
$$
\Delta(u) = -\Div(\grad u).
$$
\begin{itemize}
\item Show that, if $\Delta u \equiv 0$, then $u$ is constant.
\item Show that, if $\Delta u = \lambda u$ and $u$ is not constant, then $\lambda > 0$.
\end{itemize}

\subsection*{Problem 3}

Suppose that $f : S^n \to S^n$ is a diffeomorphism; show that $X =
D^{n+1} \cup_f D^{n+1}$ is a smooth manifold.

\subsection*{Problem 4}

Suppose $f : M^n \to \R$ is a proper smooth map for which there are no
critical points.  Is it the case that $M = N \times \R$ for some
smooth manifold $N$?

\subsection*{Problem 5}

Consider the tangent distribution on $\R^3$ given by $\ker \left( dz - y \, dx \right)$.
\begin{itemize}
\item Is this distribution involutive?
\item For any two points $p, q \in \R^3$, can you find a curve $\gamma : I
\to \R^3$ so that $\gamma'(t)$ lies in this distribution?
\end{itemize}



\end{document}
