\documentclass[12pt]{pset}

\course{Math 765}
\author{Jim Fowler}
\date{Winter 2011}

\newcommand{\CC}{\mathbb{C}}
\newcommand{\CP}{\mathbb{C}P}
\newcommand{\RP}{\mathbb{R}P}

\usepackage{enumerate}
\usepackage{nopageno}
\usepackage{hyperref}
\usepackage{tikz}

\DeclareMathOperator{\Vect}{Vect}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SO}{SO}
\DeclareMathOperator{\so}{\mathfrak{so}}
\DeclareMathOperator{\codim}{codim}
\DeclareMathOperator{\Mat}{Mat}
\DeclareMathOperator{\SymMat}{SymMat}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\Sym}{Sym}
\newcommand{\Proj}{\mathbb{P}}
\newcommand{\LL}{\mathcal{L}}

\newcommand{\HH}{H}
\newcommand{\hh}{\mathfrak{h}}

\DeclareMathOperator{\Imm}{Imm}
\DeclareMathOperator{\Mono}{Mono}
\usepackage{amsthm}
\newtheorem*{theorem*}{Theorem}

\begin{document}
\maketitle

\noindent\textbf{Immersion theory.}  This week, we'll study
$\Imm(M,N)$, the space of immersions from $M$ to $N$.  The most
important result is

\begin{theorem*}[Smale--Hirsch]
  Let $M^m$ be a closed smooth manifold, and $N^n$ a smooth manifold,
  and $m < n$.
  Then map $\Imm(M,N) \to \Mono(TM,TN)$ which sends $f$ to $Df$ is a
  homotopy equivalence.  Here, $\Mono(TM,TN)$ consists of pairs of
  continuous maps $f : M \to N$ and a smooth bundle map $TM \to
  f^{\star}TN$ which is fiberwise injective.  Note that an element of
  $\Mono(TM,TN)$ need not arise from an immersion and its derivative.
\end{theorem*}

\noindent
Email me with questions at \texttt{fowler@math.osu.edu}.  \textit{The
  exercises below should be handed in on Monday, March 7, 2011.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}[Bicycle chain redux]

  Compute $\pi_0(\Mono(TS^1, T\R^2))$, and use this result to say
  something about (regular homotopy\footnote{A \textit{regular
      homotopy} is a homotopy through immersions which extends
    continuously to a homotopy of the tangent bundles.} classes of)
  immersions of the circle in the plane.

\end{problem}

\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}[Sphere eversion is possible]

Let $i : S^2 \to \R^3$ be the standard inclusion, and let $r : S^2 \to
S^2$ be the map given by $r(v) = -v$.  Show that the immersions $i$
and $i \circ r$ are homotopic through immersions.

For a bonus point, for which $S^n \hookrightarrow \R^{n+1}$ is sphere
eversion possible?

\end{problem}

\vfill
\vfill
\vfill

\pagebreak

\null

\end{document}
