\documentclass[12pt]{pset}

\course{Math 765}
\author{Jim Fowler}
\date{Winter 2011}

\newcommand{\CC}{\mathbb{C}}
\newcommand{\CP}{\mathbb{C}P}
\newcommand{\RP}{\mathbb{R}P}

\usepackage{enumerate}
\usepackage{nopageno}
\usepackage{hyperref}

\DeclareMathOperator{\Vect}{Vect}
\DeclareMathOperator{\GL}{GL}

\begin{document}
\maketitle

\noindent\textbf{You have long understood a single vector space; now
  you will study vector spaces parametrized by a manifold---\textit{a bundle}.}
Please feel comfortable emailing me (\texttt{fowler@math.osu.edu}) if
you are having trouble with homework problems.  \textit{The exercises
  below should be handed in on Monday, January 17, 2011---but that's a
  Holiday, so let's make the deadline Wednesday, January 19, 2011,
  instead.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}[Lee 4--3]

Let $M$ be a nonempty manifold of dimension $n \geq 1$.  Show that
$\Vect(M)$ is infinite dimensional [as an $\R$-vector space].

\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}[Lee 4--9]

Let $F : M \to N$ be a local diffeomorphism.  Show that for every $Y
\in \Vect(N)$, there is a unique smooth vector field on $M$ that is
$F$-related to $Y$.

\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}
  Show that $TS^1 \cong S^1 \times \R$ and $TS^3 \cong S^3 \times
  \R^3$.
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}
Describe a nontrivial vector bundle over $S^1$.
\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{hardproblem}
Show that $TS^2 \not\cong S^2 \times \R^2$.
\end{hardproblem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}[Parallel parking is possible]

  Consider $M = S^1 \times \R^2$; a point $(\theta,x) \in M$
  corresponds to a position of the automobile $x \in \R^2$, and the
  angle $\theta$ of its wheels.

  Describe two vector fields on $M$: ``drive in direction of wheels''
  and ``turn steering wheel.''  What, in words, is the Lie bracket of
  these two vector fields?

\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}[Lee 5--5]

Let $\pi : E \to M$ and $\tilde{\pi} : \tilde{E} \to M$ be two smooth
rank-$k$ vector bundles over a smooth manifold $M$.  Suppose $\{
U_\alpha \}_{\alpha \in A}$ is an open cover of $M$ such that both $E$
and $\tilde{E}$ admit smooth local trivializations over $U_\alpha$.

Let $\{ \tau_{\alpha \beta} \}$ and $\{ \tilde{\tau}_{\alpha \beta}
\}$ denote the transition funcctions determined by the given local
trivializations of $E$ and $\tilde{E}$, respectively.

Show that $E$ and $\tilde{E}$ are smoothly isomorphic over $M$ if and
only if for each $\alpha \in A$ there exists a smooth map
$\sigma_\alpha : U_\alpha \to \GL(k,\R)$ such that
$$
\tilde{\tau}_{\alpha \beta}(p) = \sigma_\alpha(p)^{-1} \tau_{\alpha
  \beta}(p) \sigma_\beta(p), \hspace{1em} p \in U_\alpha \cap U_\beta.
$$

\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}[Lee 6--9]

Suppose $F : M \to N$ is a smooth map, $\omega$ is a smooth covector
field, and $\gamma$ a piecewise smooth curve segment in $M$.  Show
that
$$
\int_{\gamma} F^\star \omega = \int_{F \circ \gamma} \omega.
$$

\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}[Lee 6--13]

\begin{enumerate}[(a)]
\item If $M$ is a compact manifold, show that every exact covector
  field on $M$ vanishes at least at two points.
\item Is there a smooth covector field on $S^2$ that vanishes at
  exactly one point?
\end{enumerate}
\end{problem}

\end{document}
