\documentclass[12pt]{article}

\usepackage{fullpage}
\usepackage{nopageno}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\newcommand{\R}{\mathbb{R}}
\usepackage{add-copyright}

\title{Solution Set 1}
\date{Wednesday, October 1, 2008}

\begin{document}
\maketitle

\begin{enumerate}
\item For each of the following sets, find the least upper bound (if it
exists) and the greatest lower bound (if it exists).

\begin{description}
\item[(a)] $(3,5)$.\\  The least upper bound is 5.  The greatest lower bound is 3.
\vfill
\item[(b)] $[-4,17)$.\\  The least upper bound is 17.  The greatest lower bound is -4.
\vfill
\item[(c)] $[0,\infty)$.\\  There is no upper bound.  The greatest lower bound is 0.
\vfill
\item[(d)] $\{ x \in \R : x < 6 \}$.\\  The least upper bound is 6.  There is no lower bound.
\vfill
\item[(e)] $\{ x \in \R : x^2 < 2 \}$.\\  The least upper bound is $\sqrt{2}$.  The greatest lower bound is $-\sqrt{2}$.
\vfill
\item[(f)] $\{ x \in \R : |x - 1| < 3 \}$.\\  The least upper bound is $4$.  The greatest lower bound is $-2$.
\vfill
\end{description}

\item Which of the sets in Problem 1 are bounded?

Only the sets in parts \textbf{(a)}, \textbf{(b)}, \textbf{(e)}, and \textbf{(f)} are bounded.

\vfill

\item Suppose $S$ is a bounded set of real numbers, and $T$ is a
  subset of $S$.  Is $T$ also bounded?  Why or why not?

  Yes: if $S$ is bounded, and $T$ is a subset of $S$, then $T$ is
  bounded.  Why?  Because an upper (or lower) bound for $S$ is also an
  upper (or lower) bound for $T$.

  Suppose $b$ is a upper bound for $S$; by definition, this means that
  for all $s \in S$, we have $b \geq s$.

  But $T$ is a subset of $S$, so for all $t \in T$, we also have $t
  \in S$; since $b$ is bigger than (or equal to) everything in $S$, we
  have $b \geq t$.

  We conclude that for all $t \in T$, we have $b \geq t$.  This is
  what we mean when we say that $b$ is an upper bound for $T$, so $T$
  is bounded above.

  A similar argument shows that a lower bound for $S$ is also a lower bound for $T$.

\vfill

\end{enumerate}

\end{document}
