\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{nopageno}
\title{What now?  What do I do next?}

\begin{document}

\section*{What Now?  When is the next lecture?}

Well, it is over.  This is it.  There are no more lectures.  But, with
the end of calculus, your journey through mathematics has just
begun!

\subsection*{More courses to take?}

Keep taking mathematics courses!  Math 195--196 covers the mathematics
useful for economics courses; but let me also recommend the new
\textbf{Math 199}, which will focus on proofs, and prepares you for
\textbf{Math 203--205}, a year long course in analysis, after which
you can take \textbf{Math 254--256}, a year long course in algebra.

\subsection*{Other educational experiences?}

You can apply to the \textbf{Directed Reading Program}, where, if you
are accepted, you will be paired with a graduate student mentor and
work on an independent reading project.  I can write recommendation
letters for your application.

You can also apply to the department's \textbf{VIGRE REU} to do
mathematics during the summer, here in Chicago.

\subsection*{More to read?}

There are many excellent mathematics books available, depending on
your tastes.
\begin{itemize}
\item \textit{The Princeton Companion to Mathematics} by Timothy
  Gowers, June Barrow-Green, Imre Leader.  This new (2008) book is an
  amazing overview of modern mathematics.

\item \textit{The Knot Book} by Colin C. Adams.  This book is a
  lot of fun, and presents a totally different side of mathematics
  than what you may have seen thus far.

\item \textit{Generatingfunctionology} by Herbert S. Wilf.  This book
  (available on the web) introduces generating functions---a way of
  using power series to answer combinatorial questions (like ``how
  many ways can I fill a $17 \times 2$ grid with $2 \times 1$
  dominoes?'').

\item \textit{Principles of Mathematical Analysis} by Walter Rudin.
  Analysis is Calculus with all the proofs included; this serious book
  is not easy, but is the standard textbook for many analysis courses.
\end{itemize}
If you'd prefer some specific recommendations for something
interesting and different than Calculus, just let me know.


\end{document}