
\documentclass[12pt,letterpaper]{article}

\title{Syllabus}
\author{Jim Fowler}

\usepackage{multicol}
\usepackage{fullpage}
\pagestyle{empty}

\usepackage{html}

%\usepackage[T1]{fontenc}
%\usepackage{textcomp}
%\usepackage{lmodern}
%\newcommand{\ditto}{\textquotedbl}

%\setlength{\parindent}{0pt}

\newcommand{\peem}{\textsc{p.m.}}
\newcommand{\ayem}{\textsc{a.m.}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{\Large\sf Syllabus\hfill
Math 132, Section 22\hfill
Winter 2006}

Welcome to Math 132, the second quarter of a year-long Calculus
sequence.  Our goals are three-fold: study trigonometry, applications
of differentiation, and integration.

\begin{htmlonly}
You can also download this syllabus as a \htmladdnormallink{PDF
file}{http://math.uchicago.edu/~fowler/math132/syllabus.pdf}.
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Resources}

We present five resources to help you to learn Calculus.

\subsection*{Office Hours}
If you have questions, want to work through problems, or just talk
about mathematics, I invite you to come to my office hours.
\begin{multicols}{2}
\begin{tabular}{ll}
\textbf{Name:} & Jim Fowler \\
\textbf{Office:} & Math/Stat 102 \\
\textbf{Phone:} & 773--573--5659 \\
\textbf{Email:} & \texttt{jim@uchicago.edu}
\end{tabular}

\begin{tabular}{ll}
\textbf{Office Hours:} & Monday 4:30--6:30\peem \\
& Thursday 4:30--6:30\peem \\
& or by appointment
\end{tabular}
\end{multicols}
\noindent
Please email me with any concerns you have; the success of this course
depends on open communication.

\subsection*{Textbook}
Our text is \textit{Calculus} by Verberg, Purcell, and Bigdon (8th
Edition).

\subsection*{Website}
The website is on chalk; we will post homework assignments and take-home quizzes.

\subsection*{Lectures}
We meet Mondays, Wednesdays, and Fridays, 9:30--10:20\ayem\ in Ryerson
358 for an interactive lecture.

\subsection*{Problem Session}
You meet with your tutor on Tuesdays and Thursdays, 9:00--10:20\ayem\
for an interactive problem session.  The tutors, in
alphabetical order, are:
\begin{itemize}
\setlength{\itemsep}{-1ex}
\item Danny Rosenthal (\texttt{danny@uchicago.edu}) in Gates-Blake 411,
\item Sara Rezvi (\texttt{arsinoe@uchicago.edu}) in Harper 125,
\item Steve Balady (\texttt{sbalady@uchicago.edu}) in Wiebolt 106,
replacing Brian Keto.
\end{itemize}
If for any reason you want to switch to a different problem session,
contact Jim Fowler.
\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Requirements}

There are one thousand points possible in this course, broken down as follows:
\begin{description}
\item[Daily homework (100 points).]  Homework is assigned Mondays and
Wednesdays, and due the following class period.  You should work on
the homework problems together, but you must write up your solutions
independently.
\item[2 midterms (201 points each).]  The midterms are in class.  The
first midterm is on Friday, January 20, at the end of week 3; the
second midterm is on Friday, February 17, at the end of week 7.
\item[7 quizzes (14 points each).]  Short take-home quizzes are
assigned most Fridays, and due the following Monday.
\item[1 final exam (400 points).]  The final exam will be held
8:00--10:00\ayem\ on Friday, March 17, 2006.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Department Policy on Final Exams}

\textit{It is the policy of the Department of Mathematics that the
following rules apply to final exams in all undergraduate mathematics
courses:}
\begin{enumerate}
\item \textit{The final exam must occur at the time and place designated on
the College Final Exam Schedule.}  In particular, \textit{no} final examinations
may be given during the tenth week of the quarter, except in the case
of graduating seniors.
\item Any student who wishes to depart from the scheduled final exam
time for the course must receive permission from Paul Sally (office is
Ryerson 350, phone is 2-7388, email is
\texttt{sally@math.uchicago.edu}).  Instructors are not permitted to
excuse students from the scheduled time of the final exam except in
the cases of an Incomplete.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Rescheduling a Midterm}

Contact Jim Fowler early in the quarter if you will not be able to
take a midterm on the scheduled day; we likely can accomodate you, but
we will need time to prepare.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Late Quizzes and Homework}

Quite simply, you must stay caught up.  It is tempting to fall behind,
but difficult to catch up again---especially in a mathematics course.
That said, I understand your schedules are busy, so I will not
penalize you for \textit{very infrequently} turning in work a day or
two late.

\end{document}
